%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode

\documentclass[dsingle]{Dissertate}

\begin{document}

\setstretch{\dnormalspacing}
\pagenumbering{roman}
\emergencystretch 3em

\vspace{50pt}

{\Huge Corrigenda}

The publication \textit{``GARLEEK: Adding an Extra Flavor to ONIOM''} has been granted a DOI.  should reflect this change as well. Additionally, OMMProtocol's manuscript has been uploaded to the ChemRxiv pre-print server for public access. Tables 5.5 and 5.6 should reflect these changes as well. Full references would now be:

\begin{itemize}

    \item \myuline{J. Rodríguez-Guerra Pedregal}, I. Funes-Ardoiz, G. Sciortino, J.-E. Sánchez-Aparicio, G. Ujaque, A. Lledós, J.-D. Maréchal, F. Maseras. ``GARLEEK: Adding an Extra Flavor to ONIOM,'' {\em Journal of Computational Chemistry}, 2018. DOI: 10.1002/jcc.25612.

    \item \myuline{J. Rodríguez-Guerra Pedregal}, L. Alonso-Cotchico, L. Velasco-Carneros, J.-D. Maréchal. ``OMMProtocol: A command line application to launch molecular dynamics simulations with OpenMM,'' 2018. DOI: 10.26434/chemrxiv.7059263.

\end{itemize}

{\sloppy
In chapter 7, \textsc{Conclusions}, the claim on point 3 \textit{``OMMProtocol [...] has brought a 20-fold speed increase''} should cite the following benchmark: \\{\url{https://docs.google.com/spreadsheets/d/19Of_r7nHFDOyns9WhJVxDtqZ5QD5mJqWPyX2EDufDRc/view}
}

Reference on enterobactin structure [210] appears incorrectly formatted. \textit{``P. C. Database''} should be \textit{``PubChem Database''}.
\end{document}
