%!TEX root = ../dissertation.tex
% the acknowledgments section

\chapter*{Thanks}

\newthought{Getting to these last pages} means that this long, unplanned journey is getting to its end. All of it is the result of several unrelated events.

Father, you got me into science. Answering trivia on astronomy on the way to (often late) school or collecting shells of questionable origin is one of my most treasured memories. Speaking of school, teachers: Paco, Narciso, Tulio, Salvador, Carlos, Carmina, Javi. Each of you taught me more than a syllabus.

My Biotechnology classmates ---specially Bárbara, Fonti, Fer--- thanks for that conversation in our last dinner. Had not that taken place, I would not be writing this. You convinced me to follow my interests and start pivoting towards computers and programming. Enrolling in a MSc in Bioinformatics seemed like the perfect movement at the time, and today I can confirm it. Thanks Antonio for that opportunity.

Lur, I suggested you that MSc and you hit me back with a PhD. I can only hope our academic marriage lasts a few chapters more. Sonia, it is very simple: if we had not met, I would be a very different person today ---and I really like who I am right now. I cannot think of a bigger present. Lorea, neither of us enjoy this type of messages, so I'll save you the hassle. Kaixo, arreba. With no disrespect to academia, scientific progress and the public funding of research: I know the four of us have become something way bigger than everything else I might have accomplished during these last years.

Everybody else with whom I have shared a workplace: Víctor, Manu, Eli, Laia, Almudena, Pietro, Jose, Pablo Orenes, Giuseppe (oh Dioh mío, la mía nonna...), Sergi, Gantulga, Javi, Anna, Égil, Danilo, Stefano, Francesca\ldots\ We make up for the coolest Physical Chemistry department ever ---specially with the new A/C in the kitchen! I could not resist, sorry\ldots\ Blai, Pablo: thanks for make me feel at home even being 1,100 km away.  

To those students that I have mentored, even if only indirectly: Daniel Soler, Martí, Jordi, Karel, Mirea, Laura, Joe, Eric, Pablo, Daniel Viladrich, Marta, Mercè. Thanks for your contributions and ideas you brought to the table. I hope you enjoyed as much as me! 

Jean-Didier, you have taught me more things than I can count but, more importantly, I have learnt a lot of things about myself thanks to that. There is a very delicate balance sitting between trusting your gut and making rational decisions. Thanks for letting me see how that's done. I tend to be a little forgetful, but several conversations we had are now forever written in stone. Agustí, Gregori, Xavi, Jordi and all the other \textit{capos} directly or indirectly involved in this endeavour: thanks for your help, trust and support. Speaking of trust and support, thank you, Feliu and Ignacio. These last months of collaboration have been very fruitful on both personal and professional levels. 

Family, I have not been the more expressive son, brother, cousin, nephew or grandson, but I am slowly addressing that. I even feel kind of embarrased just thinking of all the things I get from you just because. I love you all.

And in a more personal note, music. Thanks for not leaving even when I looked elsewhere. We still have a long way ahead.